---
title: "Zimbabwe AEZs Map"
author: "Freedom K Muranda"
output:
  pdf_document: default
  html_notebook: default
---

\`\`\`{r setup, include=FALSE} \# Global settings for R Markdown knitr::opts_chunk\$set(echo = TRUE, fig.width = 10, fig.height = 7)

# Install and load necessary libraries

if (!requireNamespace("sf", quietly = TRUE)) install.packages("sf") if (!requireNamespace("leaflet", quietly = TRUE)) install.packages("leaflet") if (!requireNamespace("RColorBrewer", quietly = TRUE)) install.packages("RColorBrewer")

library(sf) \# For spatial data handling library(leaflet) \# For map visualization library(RColorBrewer) \# For color palettes

```{r}
# Set the path to the Zimbabwe AEZ shapefile
shapefile_path <- "/Volumes/USB DISK/Research Project 2024/New_regions/Zim_regions.shp"

# Confirm the file exists
if (!file.exists(shapefile_path)) {
  stop("Shapefile not found. Check the path and ensure all shapefile components are present.")
}

# Import the shapefile
zimbabwe_aezs <- st_read(shapefile_path)

# Validate and fix geometries, if needed
if (!all(st_is_valid(zimbabwe_aezs))) {
  zimbabwe_aezs <- st_make_valid(zimbabwe_aezs)
}

# Inspect the first few rows of the dataset
head(zimbabwe_aezs)

```

```{r}
# Define a more distinct color palette using RColorBrewer (e.g., "Paired")
color_palette <- colorFactor(palette = RColorBrewer::brewer.pal(8, "Paired"), domain = zimbabwe_aezs$Regions)

# Create a leaflet map to display Zimbabwe AEZs with rich colors and no labels
leaflet(data = zimbabwe_aezs) %>%
  addPolygons(
    fillColor = ~color_palette(Regions),  # Color palette for AEZ regions
    color = "black",                      # Black borders for polygons
    weight = 1,                           # Border weight
    opacity = 0.7,                        # Polygon opacity
    fillOpacity = 0.7,                    # Fill opacity
    popup = ~paste("Region:", Regions)    # Popup showing the region name
  ) %>%
  addLegend(
    position = "topright", 
    pal = color_palette,  # Color palette for AEZ regions
    values = zimbabwe_aezs$Regions, 
    title = "AEZ Regions",
    opacity = 1
  )  # Add the legend to the map
```
